### 2020-05-23 fig/diagram_geom.tex

\tikzset{
  auto,
  compartment/.style={
    rectangle, minimum size=9mm, rounded corners=2mm,
    thick, draw=black!15, top color=white,bottom color=black!30
  },
  %
  bigcompartment/.style={
    rectangle, minimum size=20mm, rounded corners=2mm,
    thick, draw=black!30, top color=white,bottom color=black!20
  },
  %
  point/.style={
    circle, inner sep=2pt, fill=black!5
  },
  %
  mytextbox/.style={
    rectangle, text=black!50, thin, 
    draw=white, top color=white,bottom color=white, fill=white
  }
}